%!TEX root = uw-ethesis.tex
% T I T L E   P A G E
% -------------------
% Last updated June 14, 2017, by Stephen Carr, IST-Client Services
% The title page is counted as page `i' but we need to suppress the
% page number. Also, we don't want any headers or footers.
\pagestyle{empty}
\pagenumbering{roman}

% The contents of the title page are specified in the "titlepage"
% environment.
\begin{titlepage}
        \begin{center}
        \vspace*{1.0cm}

        \Huge
        {\bf Kinodynamic Planning with $\mu$-Calculus Specifications}

        \vspace*{1.0cm}

        \normalsize
        by \\

        \vspace*{1.0cm}

        \Large
        Luc Larocque \\

        \vspace*{3.0cm}

        \normalsize
        A thesis \\
        presented to the University of Waterloo \\ 
        in fulfillment of the \\
        thesis requirement for the degree of \\
        Master of Mathematics \\
        in \\
        Applied Mathematics \\

        \vspace*{2.0cm}

        Waterloo, Ontario, Canada, 2018 \\

        \vspace*{1.0cm}

        \copyright\ Luc Larocque 2018 \\
        \end{center}
\end{titlepage}

% The rest of the front pages should contain no headers and be numbered using Roman numerals starting with `ii'
\pagestyle{plain}
\setcounter{page}{2}

\cleardoublepage % Ends the current page and causes all figures and tables that have so far appeared in the input to be printed.
% In a two-sided printing style, it also makes the next page a right-hand (odd-numbered) page, producing a blank page if necessary.

 
% E X A M I N I N G   C O M M I T T E E (Required for Ph.D. theses only)
% Remove or comment out the lines below to remove this page
\begin{center}\textbf{Examining Committee Membership}\end{center}
  \noindent
The following served on the Examining Committee for this thesis. The decision of the Examining Committee is by majority vote.
  \bigskip
  
  \noindent
\begin{tabbing}
Internal-External Member: \=  \kill % using longest text to define tab length
External Examiner: \>  Stephen Smith \\ 
\> Associate Professor, Dept.\ of Electrical and Computer Engineering \\
\end{tabbing} 
  \bigskip
  
  \noindent
\begin{tabbing}
Internal-External Member: \=  \kill % using longest text to define tab length
Supervisor: \> Jun Liu \\
\> Associate Professor, Dept.\ of Applied Mathematics \\
\end{tabbing}
  \bigskip
  
  \noindent
  \begin{tabbing}
Internal-External Member: \=  \kill % using longest text to define tab length
Internal Member: \> Brian Ingalls \\
\> Associate Professor, Dept.\ of Applied Mathematics \\
\end{tabbing}
%   \bigskip
  
%   \noindent
% \begin{tabbing}
% Internal-External Member: \=  \kill % using longest text to define tab length
% Internal-External Member: \> Deepa Thotta \\
% \> Professor, Dept. of Philosophy, University of Waterloo \\
% \end{tabbing}

\cleardoublepage

% D E C L A R A T I O N   P A G E
% -------------------------------
  % The following is a sample Delaration Page as provided by the GSO
  % December 13th, 2006.  It is designed for an electronic thesis.
  \noindent
I hereby declare that I am the sole author of this thesis. This is a true copy of the thesis, including any required final revisions, as accepted by my examiners.

  \bigskip
  
  \noindent
I understand that my thesis may be made electronically available to the public.

\cleardoublepage

% A B S T R A C T
% ---------------

\begin{center}\textbf{Abstract}\end{center}
Motion planning problems involve determining appropriate control inputs to guide a system towards a desired endpoint. Sampling-based motion planning was then developed as a technique for discretizing the state space of systems with complex environments. This makes the sampling-based method especially useful in robotics, where robots are expected to perform tasks in unknown, changing, or cluttered environments. On the other hand, temporal logic presents a means of prescribing the desired behaviour of a system. In the area of formal methods, researchers seek to solve problems in such a way that synthesized solutions provably satisfy a given temporal logic specification. In this thesis, we investigate combining the flexibility of sampling-based planning with the ability to specify the high-level behaviour of an autonomous system with the temporal logic known as \mucalc{}.

While using temporal logic specifications with motion planning has been heavily researched, reliance on an available steering function is often impractical and suited only to basic problems with linear dynamics. This is because a steering function is a solution to an optimal two-point boundary value problem (OBVP); to our knowledge, it is nearly impossible to find an analytic solution to such problems in many cases. Addressing this issue, we have developed a means of using the motion planning algorithm SST* in combination with a local model checking procedure to solve kinodynamic planning problems with deterministic \mucalc{} specifications without using a steering function. The procedure involves combining only the most pertinent information from multiple Kripke structures in order to create one abstracted Kripke structure storing the best paths to all possible proposition regions of the state-space. A linear-quadratic regulator (LQR) feedback control policy is then used to track these best paths, effectively connecting the trajectories found from multiple Kripke structures. Simulations demonstrate that it is possible to satisfy a complex liveness specification for infinitely often reaching specified regions of state-space using only forward propagation.

We proceed to repurpose this tool for real-time quadrotor motion planning with temporal logic specifications. The dynamical system is derived, and a real-time planning framework is presented based on a variant of the FMT* planning algorithm. Despite requiring a steering function, an argument is presented which allows finding OBVP solutions only for an approximation of the full dynamics. The notion of an abstracted Kripke structure is then applied in the context of quadrotor kinodynamic planning, allowing for rapid model checking and ensuring high-quality feasible solutions satisfying a given deterministic \mucalc{} specification.

\cleardoublepage

% A C K N O W L E D G E M E N T S
% -------------------------------

\begin{center}\textbf{Acknowledgments}\end{center}

Many thanks to my ever-supportive and helpful supervisor, Jun Liu. Your positivity and guidance gave me the motivation I needed to succeed, while allowing me the independence to follow the research path that suited me best. Thank you also to all of the members of the Hybrid Systems Lab. Yinan Li, Milad Farsi, Chuanzheng Wang, Riley Brooks, and Kevin Church, you have all contributed so much to my Master's experience with insightful conversation, interesting presentations, and utmost kindness.
\cleardoublepage

% D E D I C A T I O N
% -------------------

\begin{center}\textbf{Dedication}\end{center}

This is dedicated to my officially-expanded family and to Ma\v{s}a: you have made my journey through graduate studies an absolute pleasure.
\cleardoublepage

% T A B L E   O F   C O N T E N T S
% ---------------------------------
\renewcommand\contentsname{Table of Contents}
\tableofcontents
\cleardoublepage
\phantomsection    % allows hyperref to link to the correct page

% L I S T   O F   T A B L E S
% ---------------------------
% \addcontentsline{toc}{chapter}{List of Tables}
% \listoftables
% \cleardoublepage
% \phantomsection		% allows hyperref to link to the correct page

% L I S T   O F   F I G U R E S
% -----------------------------
\addcontentsline{toc}{chapter}{List of Figures}
\listoffigures
\cleardoublepage
\phantomsection		% allows hyperref to link to the correct page

% GLOSSARIES (Lists of definitions, abbreviations, symbols, etc. provided by the glossaries-extra package)
% -----------------------------
\printglossaries
\cleardoublepage
\phantomsection		% allows hyperref to link to the correct page

% Change page numbering back to Arabic numerals
\pagenumbering{arabic}

